---
title: "Weight lifting prediction"
author: "Mogens Yde-Andersen"
date: "23. aug. 2015"
output:
  html_document:
    theme: united
    toc: yes
  md_document:
    toc: yes
  word_document: default
---

github submission
gh pages på branch
###Analysis

This is my report on the practical machine learning assignment regarding the creation of a prediction model for weight lifting quality classification.

```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE)
```

Load and clean training data set

```{r "chunk1", echo=TRUE, cache=TRUE}
inTrain <- read.csv("./pml-training.csv", header=TRUE, sep=",")
inTrain <- inTrain[,-c(1:7)] ##remove experiment metadata, since they are not relevant for classe outcome
inTrain <-inTrain[,colSums(is.na(inTrain))==0] ##remove columns with missing values (since they are calculations over observations)
col_names <- c()
n <- ncol(inTrain)-1
for (i in 1:n){
  if (is.factor(inTrain[,i])){
    col_names <- c(col_names, i)
  }
} ##put all factor variables in a file - except outcome varible classe
library(lattice);library(ggplot2);library(caret)
inTrain <- inTrain[,-col_names] ##remove calculation variables and variables with no relevant data
training <- createDataPartition(y=inTrain$classe,p=0.75,list=FALSE)
trainingSet <- inTrain[training,]
testingSet <- inTrain[-training,]
```

Build a prediction model with random forest

```{r "chunk2", echo=TRUE, cache=TRUE}
library(randomForest)
modFitrf52 <- randomForest(classe~.,data=trainingSet, prox=TRUE)
modFitrf52
predrf52 <- predict(modFitrf52,newdata=testingSet[,-53])
table(testingSet$classe,predrf52)
```
Note the very low error rate of only 0.4%. This means that the random forest prediction above has an accuracy of 99.6%.

Build an alternative prediction model with classification tree (rpart).

```{r "chunk3", echo=TRUE, cache=TRUE}
library(rpart)
modFitrpart52 <- train(classe~.,data=trainingSet,method="rpart")
modFitrpart52
modFitrpart52$finalModel
predrpart52 <- predict(modFitrpart52,newdata=testingSet[,-53])
table(testingSet$classe,predrpart52)
```
Note the relatively low accuracy of 50.7%. I go with the random forest prediction model.

Cross validation on the random forest prediction model

```{r "chunk4", echo=TRUE, cache=TRUE}
seedcode <- 12345
accuracies <-c()
for (i in 1:3){
       set.seed(seedcode)
       seedcode <- seedcode+1
       training <- createDataPartition(y=inTrain$classe, p=0.75, list=FALSE)
       trainingSet<- inTrain[training,]
       testingSet<- inTrain[-training,]
       modelFitchoice <- randomForest(classe ~., data = testingSet)
       prediction <- predict(modelFitchoice, testingSet)
       testingSet$rightPred <- prediction == testingSet$classe
       t<-table(prediction, testingSet$classe)
       print(t)
       accuracy <- sum(testingSet$rightPred)/nrow(testingSet)
       print(accuracy)
}
```
I guess that the cross validation and accuracy is optimal :O).

Load and clean the testing data set

```{r "chunk5", echo=TRUE, cache=TRUE}
testing <- read.csv("./pml-testing.csv", header=TRUE, sep=",")
testing <- testing[,-c(1:7)] 
testing <-testing[,colSums(is.na(testing))==0]
predictiontesting <- predict(modelFitchoice, testing)
predictiontesting
```

